---
title: "Untitled"
author: ""
date: "March 7, 2017"
output: html_document
---




# Project Goals

The goal of this project is to work through building functions around the Jackkinfe. We will start with some simple functions and then determine a pattern to build a more general function in the end. 


## Jackknife of the Mean!

- Have an estimator $\hat{\theta}$ of parameter $\theta$  
  want the standard error of our estimate, $se_{\hat{\theta}}$
- The jackknife approximation:
    + omit case $i$, get estimate $\hat{\theta}_{(-i)}$
    + Take the variance of all the $\hat{\theta}_{(-i)}$
    + multiply that variance by $\frac{(n-1)^2}{n}$ to get $\approx$ variance of $\hat{\theta}$
- then $se_{\hat{\theta}}=$ square root of that variance


**PHP 2560 Only** (Why $(n-1)^2 / n$?  Think about just getting the standard
    error of the mean)

Answer: Because in a student t-test, we need a stardard error which takes into accound of degrees of freedom of n-1 

Write a function called `mean.jackknife` that takes argument `a_vector` and returns a jackknife estimate of the standard error.


```{r}
mean.jackknife <- function(a_vector) {
  mean_vect=a_vector
  for(i in 1:length(mean_vect)){
    mean_vect[i]=mean(a_vector[-i])
  }
  variance=var(mean_vect)
  correct=(length(mean_vect)-1)^2/(length(mean_vect))
  jackknife.stderr=sqrt(variance*correct)
  return(jackknife.stderr)
}
```



#### Jackknife for the mean Example: test your code here


```{r}
set.seed(10)
some_normals <- rnorm(1000,mean=7,sd=5)
mean(some_normals)
(formula_se_of_mean <- sd(some_normals)/sqrt(length(some_normals)))
comp=mean.jackknife(some_normals)
all.equal(formula_se_of_mean,comp)
```



##Jackknife for Gamma Parameters


The following function is a way to calculate the method of moments estimators for the gamma distribution:

```{r}
gamma.est <- function(the_data) {
  m <- mean(the_data)
  v <- var(the_data)
  a <- m^2/v
  s <- v/m
  return(c(a=a,s=s))
}
```


###Jackknife for Gamma Parameters Function

Write a function called `gamma.jackknife` that takes argument `a_vector` and returns jackknife standard error estimates on the gamma parameters. 

```{r}
gamma.jackknife <- function(a_vector) {
  as_table=matrix(0, ncol=2, nrow=length(a_vector))
  for(i in 1:length(a_vector)){
    as_table[i,]=gamma.est(a_vector[-i])
  }
    a_var=var(as_table[,1])
    s_var=var(as_table[,2])
    correct=(length(a_vector)-1)^2/(length(a_vector))
    a=sqrt(a_var*correct)
    s=sqrt(s_var*correct)
  jackknife.stderr=c(a=a,s=s)
  return(jackknife.stderr)
}
```



###Jackknife for Gamma Parameters Example


```{r}
input <- rgamma(1000, shape=0.1, scale=10)
gamma.est(input)
gamma.jackknife(input)
```



##Jackknife for linear regression coefficients

Write a function called `jackknife.lm` that takes arguments `df`, `formula` and `p` and returns jackknife standard error estimates on the coefficients of a linear regression model. 

```{r}

jackknife.lm <- function(df,formula,p) {
  
  p_table=matrix(0, ncol=p, nrow=nrow(df))
  p_table_2=matrix(0, ncol=p, nrow=1)
  for(i in 1:nrow(df)){
    new_data<-df[-i,]
    p_table[i,]<-lm(formula, new_data)$coefficients
  }

  p_table_2=apply(p_table, 2, function(x) var(x)*(nrow(df)-1)^2/(nrow(df)))
  jackknife.stderr=p_table_2
  return(jackknife.stderr)
}
```



###Jackknife for linear regression coefficients Example


```{r, eval=FALSE}
output <- 1.2 + 0.6*input +  rnorm(1000, 0, 2.1)
data <- data.frame(output,input)
my.lm <- lm(output~input, data=data)
coefficients(my.lm)
# "Official" standard errors
sqrt(diag(vcov(my.lm)))
jackknife.lm(df=data,formula="output~input",p=2)
```




##Refactoring the Jackknife


- Omitting one point or row is a common sub-task

- The general pattern:

```
figure out the size of the data
for each case
   omit that case
   repeat some estimation and get a vector of numbers
take variances across cases
scale up variances
take the square roots
```

- Refactor by extracting the common "omit one" operation

- Refactor by defining a general "jackknife" operation



###The Common Operation


- _Problem_: Omit one particular data point from a larger structure

- _Difficulty_: Do we need a comma in the index or not?

- _Solution_: Works for vectors, lists, 1D and 2D arrays, matrices, data frames:



## Goal: 


- Make the function select the correct dimensions
    - length for a 1d object
    - number of rows for 2d
- Write a function `omit.case` that omits a point given the data and returns the data minus that point. Make sure it can handle higher dimensions. 
    ```{r}
          omit.case <- function(the_data,omit=TRUE) {
            if (omit==TRUE){
              if(class(the_data)=="list" | class(the_data)=="numeric"){
                new_df=data.frame(the_data)
              }else{
                new_df=the_data
              }
            }else{
              if(class(the_data)=="list" | class(the_data)=="numeric"){
                df=data.frame(the_data)
                new_df=df[-omit,]
              }else{
                new_df=data.frame(the_data[-omit,])
              }
            }
            names(new_df)=names(the_data)
            new_df
            }
    ```
- Write a function `omit_and_est` that takes the data with an omitted point and returns whatever function your estimator does.  
    ```{r}
    
        omit_and_est <- function(the_data, estimator, omit=TRUE) {
          new_data<-omit.case(the_data, omit=TRUE)
          est_name=suppressWarnings(names(estimator(new_data)))
          if(is.na(charmatch("(Intercept)", est_name))==1){
            est=c(apply(new_data, 2, estimator))
            est_name=''
          }else{ 
            est=est.coefs(new_data)
          }
          return(est)
        }
```

```{r}
jackknife <- function(estimator,the_data, omit=TRUE) {
  
  # Define n to be lenth or number of rows here
  new_data<-omit.case(the_data, omit=TRUE)
  n=nrow(new_data)
  m=length(estimator(the_data))
  
  jackknife.ests <- matrix(sapply(1:n, omit_and_est, the_data=new_data, estimator=estimator),  ncol=n, nrow=m)
  var.of.reestimates <- apply(jackknife.ests,1,var)
  jackknife.var <- ((n-1)^2/n)* var.of.reestimates
  jackknife.stderr <- sqrt(jackknife.var)
  return(jackknife.stderr)
}
```



---

###It works


```{r}
jackknife(estimator=mean,the_data=some_normals)
all.equal(jackknife(estimator=mean,the_data=some_normals),
          mean.jackknife(some_normals))
```

---



```{r}
all.equal(jackknife(estimator=gamma.est,the_data=data$input),
          gamma.jackknife(data$input))
```

---


```{r}
all.equal(jackknife(estimator=gamma.est,the_data=data$input),
          gamma.jackknife(data$input), check.names=FALSE)
```


```{r}
est.coefs <- function(the_data) {
  return(lm(output~input,data=the_data)$coefficients)
}
est.coefs(data)
all.equal(est.coefs(data), coefficients(my.lm))
```

---


```{r}
jackknife(estimator=est.coefs,the_data=data)
all.equal(jackknife(estimator=est.coefs,the_data=data),
          jackknife.lm(df=data,formula="output~input",p=2))
```


##Further Refactoring of jackknife()


The code for `jackknife()` is still a bit clunky:
- Ugly `if-else` for finding `n`
- Bit at the end for scaling variances down to standard errors

- write a function that calculates the `n` needed for the above code:

    ```{r, eval=FALSE}
      data_size <- function(the_data) {
        return(nrow(data.frame(the_data)))
      }
    ```
- Write a function that calculate the variance of all the estimates and returns the standard error
```{r, eval=FALSE}
scale_and_sqrt_vars <- function(jackknife.ests,n) {
  var.of.reestimates <- apply(jackknife.ests,1,var)
  jackknife.var <- ((n-1)^2/n)* var.of.reestimates
  jackknife.stderr <- sqrt(jackknife.var)
  return(jackknife.stderr)
}
```



###Now invoke those functions

```{r, eval=FALSE}
jackknife <- function(estimator,the_data) {
  n <- data_size(the_data)
  jackknife.ests <- matrix(sapply(1:n, omit_and_est, the_data=the_data, estimator=estimator), ncol=n)
  return(scale_and_sqrt_vars(jackknife.ests,n))
}
```

